### 2019-07-26 appendix/appendix.Rmd
upload templates

---
title: |
       | Supplementary Materials for "PAPER"
author: 
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
- "Jessica Voigt[^voigt]"
date: July 26, 2019
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

[^freire]: Postdoctoral Research Associate, The Political Theory Project, Brown University, Providence, RI 02912, USA, <danilofreire@brown.edu>, <http://danilofreire.github.io>.

[^galdino]: Executive Diretor, Transparência Brazil, São Paulo, SP, Brazil, [mcz.fea@gmail.com](mailto:mcz.fea@gmail.com), <https://www.transparencia.org.br>.

[^mignozzetti]: School of International Relations, Fundação Getulio Vargas, São Paulo, SP, Brazil and Wilf Family Department of Politics, NYU, NY, USA, [umberto.mig@nyu.edu](mailto:umberto.mig@nyu.edu), <http://umbertomig.com>. Corresponding author.

[^voigt]: Data Scientist, Transparência Brasil, São Paulo, SP, Brazil, [voigt.jessica@gmail.com](mailto:voigt.jessica@gmail.com), <https://www.linkedin.com/in/voigtjessica>.

```{r setup, echo=FALSE}
```

# Latin American Elite Survey \label{sec:latam}

\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

Lorem ipsum.

# Bibliography