### 2019-12-19 article/letter.Rmd

---
title:
author:
date:
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: no
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)
if (!require("kableExtra")) {
    install.packages("kableExtra")
}
```

\noindent Research & Politics

\vspace{.1cm}

\noindent 17 December 2019

\vspace{.5cm}

\noindent Dear Editor and Reviewer,

\vspace{.5cm}

\noindent We would like to thank you for the opportunity to revise our
manuscript, "Bottom-Up Accountability and Public Service Provision: Evidence
from a Field Experiment in Brazil" (Manuscript ID RAP-19-0144). We have made
many revisions to the paper along the lines suggested by the editor and the
anonymous reviewer and we believe the manuscript has improved significantly as
a result.

We have worked specially hard to clarify the connection between information
provision and mobilisation, a point emphasised by both the editor and the
reviewer. Our efforts in that regard include, first, a new section in the SI
Appendix with a detailed description of _Tá de Pé_'s functionalities and the
associated Facebook campaigns. We explain the precautions we have taken to
ensure that we correctly identify the effect of the TDP app on school outcomes.
Second, we provide a thorough discussion of our additional data sources. More
specifically, we show how the Brazilian Ministry of Education collects and
publicises their data and indicate how our experiment uses that information to
evaluate the robustness of our findings. We believe the new sections help give
the results a clearer interpretation and strenghten our main arguments.

Related, we have included a discussion regarding app usage as suggested by the
reviewer. We have included information provided by Google Analytics not only on
app downloads but also on app usage, which addresses another important concern
raised by the reviewer. The data show that the app usage was significant, yet
it was not enough to trigger any responses from local governments.

Lastly, we have rewritten some parts of the main text to highlight that our
experiment hinges on participation only, and that the effect of mobilisation
and pressure on school construction outcomes require further analyses. We
believe our paper now reflects our results more accurately.

Below we discuss in more detail these and other changes we have made to the
manuscript in response to comments from the editor and the reviewer. We thank
the editor again for the opportunity to revise the manuscript, and the reviewer
for his/her extremely helpful comments.

\vspace{.5cm} \noindent Sincerely,

\vspace{.5cm} \noindent The Authors

\newpage

# Editor Comments and Responses

**1)** The editor writes: "_The reviewer confirmed my own sense that the
manuscript and underlying field experiment are interesting, although you will
see that the reviewer suggests many ways the manuscript can be improved.  I am
in full agreement with the reviewer that you need to reflect more deeply about
the theory and evidence for why information provision via the app could be
connected to mobilization, -- otherwise the null result is not interesting.  As
the reviewer notes, the null result could be due to a disconnect between
information provision and action, or due to who selects into downloading the
app._"

\vspace{.5cm} \noindent \textbf{Response:}

\begin{quotation}
We agree with the editor and the reviewer in regards to the link between
information provision and mobilisation. We believe the main issue lies in the
lack of evidence for mobilisation, that is, information on how users interacted
with the application and whether they have used the app consistently. In the
revised version of the manuscript, we provide data from Google Analytics which
indicate that users did engage with the app. According to the service, the app
had 6,092 active users in intervention 1 and gained 4,078 new users during
intervention 2. On average, each user launched 60 sessions per app install, an
indication of their actual engagement with the application. Moreover, the data
also show that the app had more than 53,000 screen visualisations, with an
average of 2.42 visualisations per user. We have added this information to the
main text and the online appendix. The additions are marked in \textbf{boldface}.
We elaborate further on this issue in the response to Reviewer 1 below.
\end{quotation}

# Reviewer 1 Comments and Responses

**1)** Reviewer 1 writes: "_My biggest concern is there is no discussion of the
actual bottom-up accountability that may (or may not) have occurred. [...] The
authors have devised a nice experiment and I believe their design to have been
appropriately administrated. But the design hinges on several critical
mechanisms that go largely undiscussed: not only must information be available
to citizens such that they might participate in local oversight, but said
citizens must 1) access the information and 2) act on said information. This is
mentioned in the Discussion---that perhaps the app was insufficient to spur
collective action---but that also seems like an empirical question the authors
could feasibly address. [...] Once people downloaded the App, did they use
it?_"

\vspace{.5cm} \noindent \textbf{Response:}

\begin{quotation}
This first point raised by the reviewer is connected to the response we have
provided above. We have made several changes to the main paper and the appendix
to address this important concern. First, we have added a the following section
to the appendix in order to clarify how the TDP app works:

\vspace{.5cm}

3. Treatment Definition and Mechanism

The \textit{Tá de Pé} app

1. Users download the app
2. Use the app and report the schools to the mayors' office
3. Our team took the reported schools and started implementing the pressure
on the mayors' office.
4. Should the pressure have work, we would witness improvements on governmental
	 data regarding the school construction progress.

For the step (4), we relied on monthly issued administrative data by the
Brazilian Ministry of Education. The Ministry of Education provides monthly
data on the school constructions that are being build with Federal resources,
and that are delayed across the country. To compute the delay we use the
finishing dates reported by the municipality at the time they sign the
agreement with the Federal Government. This data helped us to keep the App
updated and the outcomes in the paper are selected to be six months before
(pre-treatment) and after the intervention. This data was constantly updated by
the Ministry of Education, and we found no red flags in the overall quality of
the data.

Therefore, steps (1), (3), and (4) were standardized to not vary across the
country. The only source of variation in our intervention should come from step
(2).

In step (2), we applied the treatment by removing schools within municipalities
out of the app. In the first intervention, that happened at the municipal
level, we excluded all schools within a given randomly selected municipality.
In the second intervention, we excluded randomly selected schools out of the
App. Removing a school means that the school was in the dataset provided by the
Brazilian Ministry of Education, but it did not show up in the App. Therefore,
our treatment effect represents the school being present (treatment) or absent
(control) in the App.

In terms of what this represents, we argue that the App is promoting two
crucial features for bottom-up accountability: first, it is facilitating people
to connect with information that would be otherwise be in an obscure
spreadsheet at the Ministry of Education website; second, whenever people find
this information troublesome, we are providing an easy way to them to act upon
the information. As the design was intuitive and straightforward, we believe
that most of Brazilians would find it easy to use the App.

Moreover, the treatment to be successful does not require that everyone in the
municipality uses the App. Clearly that more usage means more pictures and
denounces, and this represents more pressure over the mayors office, what could
revert into a higher effect. However, just one denounce would be enough, as it
can be directed to the Brazilian General Comptroller (CGU), and this could
potentially deny future funding to the municipality.

\end{quotation}

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}