### 2019-12-20 appendix/appendix.Rmd
add appendix

---
title: |
       | Supplementary Materials for "Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil"
author:
date: \today
abstract: "Does local oversight improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app prompts users to submit data about construction sites, sends such crowdsourced information to independent engineers, and contacts the mayors' offices about project delays. Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of bottom-up monitoring on public service performance and suggest that interventions targeted to other groups, or focusing on different issues, may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

\newpage

# Balance Tests \label{sec:bal-tests}

The tables below show the results of the balance tests we conducted for
interventions 1 and 2. Our list of covariates are:

1. Log of municipal population in 2015; 
2. Log of number of poor families in each city; 
3. Log of total federal transfers to the municipality in 2016; 
4. Federal government indicator for primary school quality; 
5. Federal government indicator for secondary school quality.

The data come from the Brazilian Ministry of Education and the National Census.

In intervention 1, we randomised the treatment at the municipality level. We
selected 150 municipalities for the control group while the rest were placed in
the treatment group. The control group worked by deleting all the construction
sites in control assigned cities. No difference between control and treatment
variables is statistically significant.

```{r balance1, echo=TRUE, results='hide', warning=FALSE, message=FALSE, error=FALSE, tidy=TRUE}
# R options
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"; options(repos = r)

# Install required packages
pkgs <- c("tidyverse", "dummies", "kableExtra",
          "haven", "compareGroups", "grid", "gridExtra",
          "devtools", "lubridate", "brazilmaps",
          "abjutils", "lfe", "cobalt",
          "MatchIt", "ri2", "DeclareDesign", "knitr",
          "DiagrammeR", "DiagrammeRsvg", "rsvg",
          "flextable", "officer", "ggthemes",
          "stargazer", "formatR")

# Install required packages
install_if_not <- function(x) {
  if (x %in% rownames(installed.packages()) == FALSE)
    install.packages(x, dependencies = T)
}
lapply(pkgs, install_if_not)

# Load scripts
lapply(pkgs, require, character.only = T)
rm(pkgs, install_if_not)

# Load Data
load("../article/tdp.RData")

# Disable scientific notation
options(scipen = 999)
```

```{r balance1a, echo=TRUE, warning=FALSE, message=FALSE, error=FALSE, tidy=TRUE}
# Covariate Balance -- Intervention 1
aux <- select(impact_evaluation_phase1, treat, logPop2015, ideb_ai_2015,
              ideb_af_2015, log_poorFam2010, log_totTransf2016) %>% unique()

resu1 <- compareGroups(treat ~ logPop2015 + log_poorFam2010 +
                         log_totTransf2016 + ideb_ai_2015 + ideb_af_2015,
                       data = aux)

tab_IE1_balance <- createTable(resu1, digits = 2)
row.names(tab_IE1_balance$descr) <- c("Log Population (2015)",
                                    "Log Poor Families (2010)",
                                    "Log Total Transfers (2016)",
                                    "IDEB Initial Years (2015)",
                                    "IDEB Final Years (2015)")

print(tab_IE1_balance)
```

In intervention 2, we randomised the treatment at the school level. We placed
15% of schools that could potentially appear in the app in the control group.
We used block randomization, blocking by three characteristics:

1. Brazilian State
2. Construction Status (under construction, stopped, unfinished)
3. School construction completion rates above the national median

```{r}
## Covariate balance
aux <- select(impact_evaluation_phase2, treat, logPop2015, ideb_ai_2015,
              ideb_af_2015, log_poorFam2010, log_totTransf2016)

resu1 <- compareGroups(treat ~ logPop2015 + log_poorFam2010 +
                         log_totTransf2016 + ideb_ai_2015 + ideb_af_2015,
                       data = aux)

tab_IE2_balance <- createTable(resu1, digits = 2)
row.names(tab_IE2_balance$descr) <-  c("Log Population (2015)",
                                     "Log Poor Families (2010)",
                                     "Log Total Transfers (2016)",
                                     "IDEB Inicial Years (2015)",
                                     "IDEB Final Years (2015)")
print(tab_IE2_balance)
```

\newpage

# Manipulation Checks

In randomised experiments, manipulation is defined as the extent through which
the treatment is successfully delivered to the control and treated units. In
the case of the *Tá de Pé* app intervention, the manipulation indicates whether
the app improves school construction completion rates. We present two
manipulation indicators. First, we look into municipalities which did or did
not download the app. Next, we see the number of app downloads during
intervention 1 (February 2017--July 2018) and intervention 2 (August 2018--July
2019). The code replicates figures 2 and 3 in the article.

```{r man1, echo=TRUE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=4, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation1}Manipulation checks for intervention 1. The first plot shows the geographical distribution of the treatment condition and the second graph displays the number of TDP app downloads from August 2017 to February 2018.'}
# Map manipulation test -- Intervention 1
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase1 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$City <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase1 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 2
gaaux$Downloads <- factor(gaaux$downl, labels = c("No Downloads", "Downloads",
                                                  "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
map_downloads_int1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.001) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
map_downloads_int1

# App downloads over time -- Intervention 1
tab <- filter(ga, phase1 == 1) %>%
  group_by(data) %>%
  summarize(NewUsers = sum(novos_usuarios))
names(tab) <- c("Date", "Frequency")
freq_downloads_int1 <- ggplot(tab, aes(x = Date, y = Frequency, group = 1)) +
  geom_line() + geom_point()
freq_downloads_int1
```

```{r man2, echo=TRUE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=4, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation2}Manipulation checks for intervention 2. The graphs display the geographical and temporal variation of TDP app downloads from August 2018 to February 2019.'}
## Map manipulation test -- Intervention 2
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase2 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$City <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase2 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] = 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] = 2
gaaux$Downloads <- factor(gaaux$downl, labels = c("No Downloads", "Downloads",
                                                  "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
p1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.01) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
p1

# App downloads over time -- Intervention 2
tab <- filter(ga, phase2 == 1) %>%
  group_by(data) %>%
  summarize(NewUsers = sum(novos_usuarios))
names(tab) <- c("Date", "Frequency")
p2 <- ggplot(tab, aes(x = Date, y = Frequency, group = 1)) +
  geom_line() + geom_point()
p2
```

# Treatment Definition and Mechanism

In this section, we explain in detail how the _Tá de Pé_ app works. The
treatment involves the following procedures:

1. Users download the app.

2. Users access the app and add new information about school constructions
	 near to their location.

3. The app sends a message to the mayor's office asking details about the
	 project. If necessary, the message is also forwarded to other government
	 oversight institutions such as the city council and the Brazilian Federal
	 Comptroller. Below, we discuss the cases in which we forwarded the denounce
	 to the authorities.

4. If the pressure over the mayors has an effect, we expect to witness
	 improvements on governmental data regarding the school construction
	 progress.

The treatment is a compounded treatment comprised of steps (2) to (4). To make
sure that the only source of effect comes from step (2), we took the following
precautions:

In step (1), we have implemented Facebook campaigns in municipalities that have
school construction agreements with the Federal government. We did not
discriminate any municipality in this step, and citizens living in
municipalities in the treatment and in the control could equally have received
the banners about the _Tá de Pé_ app. The Facebook campaign was not randomised.

For step (3), we put a team of two data scientists to ensure that upon
receiving a denounce, the mayor's office, the city council, and the Brazilian
Federal Comptroller (CGU) would be notified. The decision making funnel was:

3.1. We notified the _Engineers without Borders_ (EWB) about the denounce and
sent them the user-provided pictures and auxiliary information. They issued a
report on how much the construction was likely to be delayed.

3.2. After we had the EWB report, we notified the mayor's office in compliance
with Brazil's Freedom of Information Act. We communicated that a citizen had
anonymously made a denounce about a school (fully identifying the school
construction that we were referring to), and that the Engineers Without Borders
found the construction delayed by a percentage that EWB provided. We submitted
a request for the mayor's office to provide an explanation on why the
construction was delayed, and asked for updated statistics on when the
construction will be finished. According to the FOI law, they had twenty days
to respond.

3.3. If the response was found insufficient, we forwarded the denounce to
the city council and to the Brazilian Federal Comptroller (CGU). As the
denounce goes to these institutions, they can harm mayors by restricting access
to Federal funds.

3.4. At each stage, a Twitter bot constructed by Transparência Brasil
(<https://twitter.com/tadepeapp>) reported the reception of a denounce and
tracked the responses of the mayor, the city council, and the CGU.

In step (4), we relied on administrative data by the Brazilian Ministry of
Education. The Ministry of Education provides monthly data on school
constructions build with Federal Government resources and shows how many of
them are delayed across the country. We use the finishing dates reported by the
municipality at the time they sign the agreement with the Federal Government as
our measure of construction delays. This data helped us to keep the app updated
and the outcomes in the paper are selected to be six months before
(pre-treatment) and after the intervention. This data was constantly updated by
the Ministry of Education, and we found no red flags in the overall quality of
the data.

Therefore, steps (1), (3), and (4) were standardised and do not vary across the
country. The only source of variation in our intervention should come from step
(2).

In step (2), we applied the treatment by removing schools within municipalities
out of the app. In the first intervention, which happened at the municipal
level, we excluded all schools within a given randomly selected municipality.
In the second intervention, we excluded randomly selected schools out of the
app. Removing a school means that the school was in the dataset provided by the
Brazilian Ministry of Education, but it did not show up in the app. Therefore,
our treatment effect represents the school being present (treatment) or absent
(control) in the app.

We argue that the app promotes two crucial features of bottom-up
accountability. First, it helps people to access information that would
otherwise be available only in an obscure spreadsheet on the Ministry of
Education website. Second, whenever people find this information troublesome,
we are providing an easy way to them to act upon the information. As the _Tá de
Pé_ design was intuitive and straightforward, we believe that most of
Brazilians would find it easy to use the app.

Moreover, the success of treatment does not require that everyone in the
municipality uses it. The more usage means more pictures and denounces, and
this represents more pressure over the mayors office, what could revert into a
higher effect. However, just one denounce would be enough for the project to be
successful, as the denounce can be directed to the Brazilian General
Comptroller (CGU), and this could potentially deny future funding to the
municipality.

# Outcome Data Collection

We collect the main outcomes using the data reported by the municipality to the
Brazilian Ministry of Education. Note that this is the same data as the
pre-treatment data that we use to feed in the app with delayed schools.

The outcomes used in the paper are:

1. **Percentage of the project completed before the impact evaluation
	 started**: This is a column in the Ministry of Education dataset. The data
	 in this column were sampled in the beginning of the intervention.

2. **Percentage of the project reported as completed by the end of the
	 intervention period**: Another column in the Ministry of Education
	 dataset. The data  were sampled six months after the intervention.

3. **Difference between the percentage reported as completed before and after
	 the intervention**: The difference between (1) and (2) above. Note
	 that according to @gerber2012field, this way to look into outcomes is
	 positive, as having a pre-treatment covariate that highly predicts the
	 outcome reduces the variance in the estimator considerably.

4. **Dummy indicator for finished constructions**: The Ministry of Education
	 dataset has a column with the construction status. We created a dummy to be
	 one when the construction was reported as finished and zero otherwise. The
	 data here were sampled six months after the intervention.

5. **Dummy indicator for cancelled constructions**: Similarly, we created a
	dummy variable from the Ministry of Education dataset indicating when the
	construction was reported as cancelled. The data here were sampled six months
	after the intervention.

6. **Number of schools where construction companies updated the conclusion
	 dates**: Lastly, the Ministry of Education dataset has a column with the
	 construction estimated completion date. We took the completion dated
	 reported before and six months after the intervention. If the dates
	 differed, we coded as one, and zero otherwise.

All these variables come from our manipulation of the Ministry of Education
data. Collecting the outcomes in this way is superior than the alternative of
gathering the data ourselves. First, we would not be able to manipulate the
results and cherry-pick favourable results. Even if they have the best
intentions, researchers may bias the results just by the choices they make in
the data collection process. For instance, if we send a RA to collect the data,
as she knows the app, she could unconsciously have a tendency to over-report
positive results. This would severely bias our findings and undermine the
credibility of our experiment. Second, using external data makes the results
easily reproducible by other researchers. As the Ministry of Education has
standardised data collection procedures, they greatly simplify and facilitate
the reproduction of our findings.

# Facebook Tá de Pé Campaign

The app by itself would probably not have had an effect on the responsiveness
of mayors. To boost usage and publicize the app, Transparência Brasil ran 64
Facebook campaigns throughout the country. The Facebook campaign concentrates
the bulk of the investments in the beginning of the intervention. The app
download data is in the second panel of Figures 2 and 3. For the Facebook
interaction, we have the following statistics:

1. Impressions
  - Mean: `r mean(fballt$impressoes)`
  - Total for all campaigns: `r sum(fballt$impressoes)`

2. App Download link clicks:
  - Mean: `r mean(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular'])`
  - Total for all campaigns: `r sum(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular'])`

As we see from the actual download activity, the people who indeed downloaded
the app were much fewer than these numbers would suggest.

We have chosen Facebook because the platform offers an easy and cheap method to
spread information. Moreover, according to the Statista website, Facebook is
used by 70 million Brazilians (60% of the entire adult population), and the
demographics largely match the demographics publicised by the Brazilian
Institute of Geography and Statistics. For instance, see the following
webpages:

1. [Statista Total Facebook Users](https://www.statista.com/statistics/244936/number-of-facebook-users-in-brazil/)
2. [Statista Total Facebook Users by Age](https://www.statista.com/statistics/866282/facebook-user-share-brazil-age/)

\newpage

# Main Results

## Intervention 1

The code below replicates the findings of table 1 in the manuscript.

```{r int1, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
load('../article/tdp.RData')

# Load packages
pkgs <- c('tidyverse', 'dummies', 'ggpubr',
          'haven', 'compareGroups',
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt',
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg',
          'flextable', 'officer', 'ggthemes')
invisible(lapply(pkgs, require, character.only = T))

# Auxiliary function: calculate coefficients
calcATEs <- function(outcs, treats, ctrls, fes, cls, dat, wts = NULL, ...) {
  ctrls_mod <- paste(ctrls, collapse = '+')
  fes_mod <- paste(fes, collapse = '+')
  moddescr <- data.frame()
  mods <- list()
  dataux <- dat
  k <- 1
  for (i in treats) {
    for (j in outcs) {
      if(is.null(wts)) {
        dat <- dataux %>%
          select(one_of(c(j, treats, ctrls, fes, cls))) %>%
          na.omit()
        wtsaux <- NULL
      } else {
          dat <- dataux %>%
            select(one_of(c(j, treats, ctrls, fes, cls, wts))) %>%
            na.omit()
          wtsaux <- dat[,wts]
      }
      # No controls and no fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(i,paste('',0,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = F))
      print(k); k = k+1
      # Controls and no fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i,ctrls_mod), collapse = '+'),
                        paste('',0,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = F))
      print(k); k = k+1
      # No controls and fes
      mods[[k]] <- felm(
        as.formula(paste(c(j,paste(paste(c(i), collapse = '+'),
                                   paste('',fes_mod,0,cls, sep = '|'),
                                   sep = '')),
                         collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = T))
      print(k); k = k+1
      # Controls and fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i,ctrls_mod), collapse = '+'),
                        paste('',fes_mod,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = T))
      print(k); k = k+1
    }
  }
  return(list(moddescr, mods))
}

####
## Analysis Intervention 1
####

## ATEs
outcs <- c('percExecBegin', 'percExecEnd',
           'deltaPercExec', 'isFinished',
           'isCancelled', 'updDate')
treats <- c('treat')
ctrls <- c('logPop2015', 'ideb_ai_2015','ideb_af_2015',
           'log_poorFam2010', 'log_totTransf2016')
fes <- c('state')
cls <- c('municipality')

resInt1 <- calcATEs(outcs, treats, ctrls, fes, cls, dat = impact_evaluation_phase1)

# Results intervention 1
stargazer(resInt1[[2]][[4]],resInt1[[2]][[8]],resInt1[[2]][[12]],
          resInt1[[2]][[16]],resInt1[[2]][[20]],resInt1[[2]][[24]],
          title = '\\label{tab:results1} Impact Evaluation -- Intervention 1',
          keep='treatTreatment',
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = 'Cluster-robust SEs at the municipality level.',
          add.lines = list(c("Controls", "Yes", "Yes", "Yes",
                             "Yes", "Yes", "Yes"),
            c("State fixed effects", "Yes", "Yes", "Yes",
              "Yes", "Yes", "Yes")),
          notes.append = T, header=FALSE)

# Results intervention 1 -- No Controls and No Fixed Effects
stargazer(resInt1[[2]][[1]],resInt1[[2]][[5]],resInt1[[2]][[9]],
          resInt1[[2]][[13]],resInt1[[2]][[17]],resInt1[[2]][[21]],
          title = '\\label{tab:results1} Impact Evaluation -- Intervention 1 \\\\ No Controls and No Fixed Effects',
          keep='treatTreatment',
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = 'Cluster-robust SEs at the municipality level.',
          add.lines = list(c("Controls", "No", "No", "No",
                             "No", "No", "No"),
            c("State fixed effects", "No", "No", "No",
              "No", "No", "No")),
          notes.append = T, header=FALSE)

# Results intervention 1 -- No Controls
stargazer(resInt1[[2]][[3]],resInt1[[2]][[7]],resInt1[[2]][[11]],
          resInt1[[2]][[15]],resInt1[[2]][[19]],resInt1[[2]][[23]],
          title = '\\label{tab:results1} Impact Evaluation -- Intervention 1 \\\\ No Controls',
          keep='treatTreatment',
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = 'Cluster-robust SEs at the municipality level.',
          add.lines = list(c("Controls", "No", "No", "No",
                             "No", "No", "No"),
            c("State fixed effects", "Yes", "Yes", "Yes",
              "Yes", "Yes", "Yes")),
          notes.append = T, header=FALSE)

# Results intervention 1 -- No Fixed Effects
stargazer(resInt1[[2]][[2]],resInt1[[2]][[6]],resInt1[[2]][[10]],
          resInt1[[2]][[14]],resInt1[[2]][[18]],resInt1[[2]][[22]],
          title = '\\label{tab:results1} Impact Evaluation -- Intervention 1 \\\\ No Fixed Effects',
          keep='treatTreatment',
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = 'Cluster-robust SEs at the municipality level.',
          add.lines = list(c("Controls", "Yes", "Yes", "Yes",
                             "Yes", "Yes", "Yes"),
            c("State fixed effects", "No", "No", "No",
              "No", "No", "No")),
          notes.append = T, header=FALSE)

```

\vspace{.5cm}

\begin{table}[!htbp] \centering
  \caption{\label{tab:results1} Impact Evaluation -- Intervention 1}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
 ATE & $-$0.99 & $-$1.12 & $-$0.13 & 0.002 & 0.02$^{**}$ & 0.05 \\
  & (2.59) & (2.96) & (1.03) & (0.01) & (0.01) & (0.05) \\
  & & & & & & \\
\hline \\[-1.8ex]
Controls & Yes & Yes & Yes & Yes & Yes & Yes \\
State fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\
Observations & 2,986 & 2,986 & 2,986 & 2,986 & 2,986 & 2,926 \\
R$^{2}$ & 0.16 & 0.14 & 0.05 & 0.02 & 0.07 & 0.13 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
\end{tabular}
\end{table}

The models without control variables or fixed effects follow below.

\begin{table}[!htbp] \centering
  \caption{\label{tab:results1} Impact Evaluation -- Intervention 1 \\ No Controls and No Fixed Effects}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
 ATE & $-$0.21 & $-$0.56 & $-$0.36 & 0.01 & 0.02$^{**}$ & 0.04 \\
  & (4.10) & (4.86) & (1.21) & (0.01) & (0.01) & (0.07) \\
  & & & & & & \\
\hline \\[-1.8ex]
Controls & No & No & No & No & No & No \\
State fixed effects & No & No & No & No & No & No \\
Observations & 2,986 & 2,986 & 2,986 & 2,986 & 2,986 & 2,926 \\
R$^{2}$ & 0.0000 & 0.0000 & 0.0001 & 0.0001 & 0.001 & 0.001 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
\end{tabular}
\end{table}

\begin{table}[!htbp] \centering
  \caption{\label{tab:results1} Impact Evaluation -- Intervention 1 \\ No Controls}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
 ATE & $-$2.20 & $-$2.08 & 0.12 & 0.003 & 0.03$^{**}$ & 0.03 \\
  & (2.87) & (3.33) & (1.06) & (0.01) & (0.01) & (0.05) \\
  & & & & & & \\
\hline \\[-1.8ex]
Controls & No & No & No & No & No & No \\
State fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\
Observations & 2,986 & 2,986 & 2,986 & 2,986 & 2,986 & 2,926 \\
R$^{2}$ & 0.07 & 0.06 & 0.04 & 0.01 & 0.05 & 0.06 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
\end{tabular}
\end{table}

\begin{table}[!htbp] \centering
  \caption{\label{tab:results1} Impact Evaluation -- Intervention 1 \\ No Fixed Effects}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
 ATE & 1.23 & 0.82 & $-$0.41 & 0.01 & 0.02$^{**}$ & 0.06 \\
  & (3.59) & (4.14) & (1.06) & (0.01) & (0.01) & (0.06) \\
  & & & & & & \\
\hline \\[-1.8ex]
Controls & Yes & Yes & Yes & Yes & Yes & Yes \\
State fixed effects & No & No & No & No & No & No \\
Observations & 2,986 & 2,986 & 2,986 & 2,986 & 2,986 & 2,926 \\
R$^{2}$ & 0.11 & 0.10 & 0.02 & 0.01 & 0.02 & 0.08 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
\end{tabular}
\end{table}

\newpage

## Intervention 2

Below, the code required to replicate the main findings of intervention 2.

```{r int2, echo=FALSE, warning=FALSE, message=FALSE, results='hide',cache=TRUE}
load('../article/tdp.RData')

# Load packages
pkgs <- c('tidyverse', 'dummies', 'ggpubr',
          'haven', 'compareGroups',
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt',
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg',
          'flextable', 'officer', 'ggthemes',
          'stargazer')
invisible(lapply(pkgs, require, character.only = T))

# Auxiliary function: calculate coefficients
calcATEs <- function(outcs, treats, ctrls, fes, cls, dat, wts = NULL, ...) {
  ctrls_mod <- paste(ctrls, collapse = '+')
  fes_mod <- paste(fes, collapse = '+')
  moddescr <- data.frame()
  mods <- list()
  dataux <- dat
  k <- 1
  for (i in treats) {
    for (j in outcs) {
      if(is.null(wts)) {
        dat <- dataux %>%
          select(one_of(c(j, treats, ctrls, fes, cls))) %>%
          na.omit()
        wtsaux <- NULL
      } else {
          dat <- dataux %>%
            select(one_of(c(j, treats, ctrls, fes, cls, wts))) %>%
            na.omit()
          wtsaux <- dat[,wts]
      }
      # No controls and no fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(i,paste('',0,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = F))
      print(k); k = k+1
      # Controls and no fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i,ctrls_mod), collapse = '+'),
                        paste('',0,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = F))
      print(k); k = k+1
      # No controls and fes
      mods[[k]] <- felm(
        as.formula(paste(c(j,paste(paste(c(i), collapse = '+'),
                                   paste('',fes_mod,0,cls, sep = '|'),
                                   sep = '')),
                         collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = F,
                                   hFEs = T))
      print(k); k = k+1
      # Controls and fes
      mods[[k]] <- felm(as.formula(
        paste(c(j,paste(paste(c(i,ctrls_mod), collapse = '+'),
                        paste('',fes_mod,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
      moddescr <- rbind(moddescr,
                        data.frame(outc = j,
                                   treat = i,
                                   hCtrls = T,
                                   hFEs = T))
      print(k); k = k+1
    }
  }
  return(list(moddescr, mods))
}

outcs <- c('percExecBegin', 'percExecEnd',
           'deltaPercExec', 'isFinished',
           'isCancelled', 'updDate')
ctrls <- c('logPop2015', 'ideb_ai_2015','ideb_af_2015',
           'log_poorFam2010', 'log_totTransf2016')
fes <- c('state')
cls <- c('municipality')

resInt2 <- calcATEs(outcs, 'treat', ctrls, fes, cls, dat = impact_evaluation_phase2, wts = 'IPW_treat')

# Results intervention 2
library(stargazer)
stargazer(resInt2[[2]][[4]],resInt2[[2]][[8]],resInt2[[2]][[12]],
          resInt2[[2]][[16]],resInt2[[2]][[20]],resInt2[[2]][[24]],
          title = '\\label{tab:results2} Impact Evaluation -- Intervention 2',
          dep.var.labels.include=F,
          keep='treatAppTreatment', type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = c('Cluster-robust SEs at the municipality level.', 'IPW computed using the randomizr package'),
          add.lines = list(c("Controls", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes"),
            c("State Fixed Effects", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes")),
          notes.append = T, header=FALSE)

# Results intervention 2 -- No Controls and No Fixed Effects
stargazer(resInt2[[2]][[1]],resInt2[[2]][[5]],resInt2[[2]][[9]],
          resInt2[[2]][[13]],resInt2[[2]][[17]],resInt2[[2]][[21]],
          title = '\\label{tab:results2} Impact Evaluation -- Intervention 2 \\\\ No Controls and No Fixed Effects',
          keep='treatTreatment',
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = c('Cluster-robust SEs at the municipality level.', 'IPW computed using the randomizr package'),
          add.lines = list(c("Controls", "No", "No", "No",
                             "No", "No", "No"),
            c("State fixed effects", "No", "No", "No",
              "No", "No", "No")),
          notes.append = T, header=FALSE)

# Results intervention 1 -- No Controls
stargazer(resInt2[[2]][[3]],resInt2[[2]][[7]],resInt2[[2]][[11]],
          resInt2[[2]][[15]],resInt2[[2]][[19]],resInt2[[2]][[23]],
          title = '\\label{tab:results2} Impact Evaluation -- Intervention 2 \\\\ No Controls',
          keep='treatTreatment',
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = c('Cluster-robust SEs at the municipality level.', 'IPW computed using the randomizr package'),
          add.lines = list(c("Controls", "No", "No", "No",
                             "No", "No", "No"),
            c("State fixed effects", "Yes", "Yes", "Yes",
              "Yes", "Yes", "Yes")),
          notes.append = T, header=FALSE)

# Results intervention 1 -- No Fixed Effects
stargazer(resInt2[[2]][[2]],resInt2[[2]][[6]],resInt2[[2]][[10]],
          resInt2[[2]][[14]],resInt2[[2]][[18]],resInt2[[2]][[22]],
          title = '\\label{tab:results2} Impact Evaluation -- Intervention 2 \\\\ No Fixed Effects',
          keep='treatTreatment',
          type = 'latex',
          covariate.labels = 'ATE',
          column.labels = c("Investment & Investment & Delta & Finished & Cancelled & Updated \\\\ & Before & After & Investment & Construction & Construction & Date \\\\"),
          dep.var.labels.include=F, column.sep.width = '-2pt',
          omit.stat=c('adj.rsq', 'ser'), align=F, digits=2,
          notes = c('Cluster-robust SEs at the municipality level.', 'IPW computed using the randomizr package'),
          add.lines = list(c("Controls", "Yes", "Yes", "Yes",
                             "Yes", "Yes", "Yes"),
            c("State fixed effects", "No", "No", "No",
              "No", "No", "No")),
          notes.append = T, header=FALSE)
```

\vspace{.5cm}

\begin{table}[!htbp] \centering
  \caption{\label{tab:results2} Impact Evaluation -- Intervention 2}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
\hline \\[-1.8ex]
Controls & Yes & Yes & Yes & Yes & Yes & Yes \\
State Fixed Effects & Yes & Yes & Yes & Yes & Yes & Yes \\
Observations & 3,226 & 3,226 & 3,226 & 3,226 & 3,226 & 3,109 \\
R$^{2}$ & 0.12 & 0.11 & 0.04 & 0.02 & 0.17 & 0.09 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
 & \multicolumn{6}{r}{IPW computed using the randomizr package} \\
\end{tabular}
\end{table}

And the models without controls or fixed effects follow below.

\begin{table}[!htbp] \centering
  \caption{\label{tab:results2} Impact Evaluation -- Intervention 2 \\ No Controls and No Fixed Effects}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
 ATE & $-$1.02 & $-$1.90 & $-$0.88 & $-$0.001 & 0.01 & $-$0.002 \\
  & (1.52) & (1.66) & (0.61) & (0.01) & (0.02) & (0.02) \\
  & & & & & & \\
\hline \\[-1.8ex]
Controls & No & No & No & No & No & No \\
State fixed effects & No & No & No & No & No & No \\
Observations & 3,226 & 3,226 & 3,226 & 3,226 & 3,226 & 3,109 \\
R$^{2}$ & 0.0003 & 0.001 & 0.001 & 0.0000 & 0.0004 & 0.0000 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
 & \multicolumn{6}{r}{IPW computed using the randomizr package} \\
\end{tabular}
\end{table}

\begin{table}[!htbp] \centering
  \caption{\label{tab:results2} Impact Evaluation -- Intervention 2 \\ No Controls}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
 ATE & $-$1.46 & $-$2.30 & $-$0.84 & $-$0.002 & 0.02 & $-$0.003 \\
  & (1.49) & (1.62) & (0.61) & (0.01) & (0.02) & (0.02) \\
  & & & & & & \\
\hline \\[-1.8ex]
Controls & No & No & No & No & No & No \\
State fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\
Observations & 3,226 & 3,226 & 3,226 & 3,226 & 3,226 & 3,109 \\
R$^{2}$ & 0.05 & 0.06 & 0.03 & 0.01 & 0.09 & 0.05 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
 & \multicolumn{6}{r}{IPW computed using the randomizr package} \\
\end{tabular}
\end{table}

\begin{table}[!htbp] \centering
  \caption{\label{tab:results2} Impact Evaluation -- Intervention 2 \\ No Fixed Effects}
  \label{}
\begin{tabular}{@{\extracolsep{-2pt}}lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\
\cline{2-7}
 & Investment & Investment & Delta & Finished & Cancelled & Updated \\ & Before & After & Investment & Construction & Construction & Date \\ &  &  &  &  &  \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
\hline \\[-1.8ex]
 ATE & $-$0.94 & $-$1.90 & $-$0.96 & $-$0.0005 & 0.01 & $-$0.002 \\
  & (1.47) & (1.59) & (0.61) & (0.01) & (0.01) & (0.02) \\
  & & & & & & \\
\hline \\[-1.8ex]
Controls & Yes & Yes & Yes & Yes & Yes & Yes \\
State fixed effects & No & No & No & No & No & No \\
Observations & 3,226 & 3,226 & 3,226 & 3,226 & 3,226 & 3,109 \\
R$^{2}$ & 0.08 & 0.08 & 0.01 & 0.005 & 0.12 & 0.05 \\
\hline
\hline \\[-1.8ex]
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
 & \multicolumn{6}{r}{Cluster-robust SEs at the municipality level.} \\
 & \multicolumn{6}{r}{IPW computed using the randomizr package} \\
\end{tabular}
\end{table}

\newpage

# Randomisation Inference

We employ randomised inference to estimate the probability of the sharp null hypothesis over all possible randomisations that could have occurred under our research design. We fail to reject the null in all but the finished school indicator in experiment 1.

```{r ri, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=TRUE, fig.width=7, fig.height=8, fig.show='hold', fig.align='center', fig.cap='\\label{fig:randomisation-inference}Sampling distribution of the estimated coefficient for our six outcomes in two interventions. Graphs on the left correspond to randomisation inference estimates for intervention 1 and those on the right describe the results for intervention 2.'}
# Chunk randomization inference
load('../article/tdp.RData')

# Load packages
pkgs <- c('tidyverse', 'dummies',
          'haven', 'compareGroups',
          'devtools', 'lubridate', 'brazilmaps',
          'abjutils', 'lfe', 'cobalt',
          'MatchIt', 'ri2', 'DeclareDesign', 'knitr',
          'DiagrammeR', 'DiagrammeRsvg', 'rsvg',
          'flextable', 'officer', 'ggthemes', 'ggpubr')
invisible(lapply(pkgs, require, character.only = T))

# Function for custom randomization inference
test_fun <- function(dat, outc, treat, wts = NULL) {
  ctrls_mod <- paste(ctrls, collapse = '+')
  fes_mod <- paste(fes, collapse = '+')
  moddescr <- data.frame()
  mods <- list()
  if(is.null(wts)) {
    dat <- dat %>%
      select(one_of(c(outc, treat, ctrls, fes, cls))) %>%
      na.omit()
    wtsaux <- NULL
  } else {
    dat <- dat %>%
      select(one_of(c(outc, treat, ctrls, fes, cls, wts))) %>%
      na.omit()
    wtsaux <- dat[,wts]
  }
  aux <-
    felm(as.formula(
      paste(c(outc,paste(paste(c(treat,ctrls_mod), collapse = '+'),
                        paste('',fes_mod,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
  return(aux$coefficients[1])
}

outcs <- c('percExecBegin', 'percExecEnd',
           'deltaPercExec', 'isFinished',
           'isCancelled', 'updDate')
ctrls <- c('logPop2015', 'ideb_ai_2015','ideb_af_2015',
           'log_poorFam2010', 'log_totTransf2016')
fes <- c('state')
cls <- c('municipality')

# First experiment
set.seed(543129) #Random.org
auxdat <- impact_evaluation_phase1[,c('ibge_code7', outcs, ctrls, fes, cls)]
aux2 <- unique(impact_evaluation_phase1[,c('ibge_code7', 'treat')]) %>% na.omit()
N = sum(table(aux2$treat))
Ctrl = 150
resri1 <- data.frame(1:1000)
for (k in 1:1000) {
    treat <- c(rep(1,N-Ctrl), rep(0,Ctrl))
    treat <- sample(treat, N)
    aux2 <- data.frame(aux2, treat = treat)
}
auxdat <- left_join(auxdat, aux2)
for (i in outcs) {
  res <- numeric()
  for (k in 1:1000)
    res[k] <- test_fun(auxdat, i, paste('treat', k, sep = '.'))
  resri1 <- data.frame(resri1, res)
  names(resri1)[names(resri1)=='res'] = i
}
resri1 <- resri1[,-1]
i1h1 <- ggplot(data=resri1, aes(resri1$percExecBegin)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase1, 'percExecBegin', 'treat'),
             lwd = 1, lty = 2) +
  labs(x="Percentage Completed Before\n(placebo)", y="Frequency")
i1h2 <- ggplot(data=resri1, aes(resri1$percExecEnd)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase1, 'percExecEnd', 'treat'),
             lwd = 1, lty = 2) +
  labs(x="Percentage Completed End Intervention", y="Frequency")
i1h3 <- ggplot(data=resri1, aes(resri1$deltaPercExec)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase1, 'deltaPercExec', 'treat'),
             lwd = 1, lty = 2) +
  labs(x="Diff. Pct. Completed Before and After", y="Frequency")
i1h4 <- ggplot(data=resri1, aes(resri1$isFinished)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase1, 'isFinished', 'treat'),
             lwd = 1, lty = 2) +
  labs(x="Finished School Construction Indicator", y="Frequency")
i1h5 <- ggplot(data=resri1, aes(resri1$isCancelled)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase1, 'isCancelled', 'treat'),
             lwd = 1, lty = 2) +
  labs(x="Cancelled School Construction Indicator", y="Frequency")
i1h6 <- ggplot(data=resri1, aes(resri1$updDate)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase1, 'updDate', 'treat'),
             lwd = 1, lty = 2) +
  labs(x="Updated Endline Date School Construction", y="Frequency")

## Intervention 2
# Declare RA
aux <- filter(rand2018, !is.na(Z_intheapp))
declRA <- declare_ra(blocks = aux$var_blocking,
                     prob_each = c(.85,.15),
                     conditions = c(1, 0))

# Function for custom randomization inference
test_fun <- function(dat, outc, treat, wts = NULL) {
  ctrls_mod <- paste(ctrls, collapse = '+')
  fes_mod <- paste(fes, collapse = '+')
  moddescr <- data.frame()
  mods <- list()
  if(is.null(wts)) {
    dat <- dat %>%
      select(one_of(c(outc, treat, ctrls, fes, cls))) %>%
      na.omit()
    wtsaux <- NULL
  } else {
    dat <- dat %>%
      select(one_of(c(outc, treat, ctrls, fes, cls, wts))) %>%
      na.omit()
    wtsaux <- dat[,wts]
  }
  aux <-
    felm(as.formula(
      paste(c(outc,paste(paste(c(treat,ctrls_mod), collapse = '+'),
                        paste('',fes_mod,0,cls, sep = '|'), sep = '')),
              collapse = '~')), data = dat, weights = wtsaux)
  return(aux$coefficients[1])
}

outcs <- c('percExecBegin', 'percExecEnd',
           'deltaPercExec', 'isFinished',
           'isCancelled', 'updDate')
ctrls <- c('logPop2015', 'ideb_ai_2015','ideb_af_2015',
           'log_poorFam2010', 'log_totTransf2016')
fes <- c('state')
cls <- c('municipality')

# First experiment
set.seed(543129) #Random.org
auxdat <- impact_evaluation_phase2[,c('id_project', outcs, ctrls, fes, cls)]
aux2 <- unique(aux[,c('id', 'Z_intheapp', 'IPW_intheapp')]) %>%
  na.omit() %>%
  rename(id_project = id,
         treat = Z_intheapp,
         wts = IPW_intheapp) %>%
  mutate(treat = as.numeric(as.factor(treat))-1)
resri2 <- data.frame(1:1000)
for (k in 1:1000) {
    Z <- block_ra(blocks = aux$var_blocking,
                  conditions = c(1, 0),
                  prob_each = c(.85,.15))
    IPW <- 1/obtain_condition_probabilities(declRA, Z)
    aux2 <- data.frame(aux2, treat = Z, wts = IPW)
}
auxdat <- left_join(auxdat, aux2)
for (i in outcs) {
  res <- numeric()
  for (k in 1:1000)
    res[k] <- test_fun(auxdat, i, paste('treat', k, sep = '.'),
                       wts = paste('wts', k, sep = '.'))
  resri2 <- data.frame(resri2, res)
  names(resri2)[names(resri2)=='res'] = i
}
resri2 <- resri2[,-1]
i2h1 <- ggplot(data=resri2, aes(resri2$percExecBegin)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase2,
                                   'percExecBegin', 'treat', wts = 'IPW_treat'),
             lwd = 1, lty = 2) +
  labs(x="Percentage Completed Before\n(placebo)", y="Frequency")
i2h2 <- ggplot(data=resri2, aes(resri2$percExecEnd)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase2,
                                   'percExecEnd', 'treat', wts = 'IPW_treat'),
             lwd = 1, lty = 2) +
  labs(x="Percentage Completed End Intervention", y="Frequency")
i2h3 <- ggplot(data=resri2, aes(resri2$deltaPercExec)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase2,
                                   'deltaPercExec', 'treat', wts = 'IPW_treat'),
             lwd = 1, lty = 2) +
  labs(x="Diff. Pct. Completed Before and After", y="Frequency")
i2h4 <- ggplot(data=resri2, aes(resri2$isFinished)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase2,
                                   'isFinished', 'treat', wts = 'IPW_treat'),
             lwd = 1, lty = 2) +
  labs(x="Finished School Construction Indicator", y="Frequency")
i2h5 <- ggplot(data=resri2, aes(resri2$isCancelled)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase2,
                                   'isCancelled', 'treat', wts = 'IPW_treat'),
             lwd = 1, lty = 2) +
  labs(x="Cancelled School Construction Indicator", y="Frequency")
i2h6 <- ggplot(data=resri2, aes(resri2$updDate)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = test_fun(impact_evaluation_phase2,
                                   'updDate', 'treat', wts = 'IPW_treat'),
             lwd = 1, lty = 2) +
  labs(x="Updated Endline Date School Construction", y="Frequency")

ggarrange(i1h1, i2h1,
          i1h2, i2h2,
          i1h3, i2h3,
          i1h4, i2h4,
          i1h5, i2h5,
          i1h6, i2h6,
          labels = rep(c("Int. 1", "Int. 2"), 6),
          label.x = .16, label.y = .9,
          font.label = list(size = 12, color = "black", face = "plain"),
          ncol = 2, nrow = 6)
```

\newpage

# APSA Experimental Section Standard Report for Experimental Research

## Hypotheses

**Hypotheses**: The experiment studies how technology facilitates bottom-up and
top-down pressure to improve service provision. More precisely, we test whether
the *Tá de Pé* mobile phone app is able to successfully mobilise citizens to
exert pressure on local representatives and civil servants. We also test
whether bottom-up monitoring of public works improves school construction rates
and spending.

\noindent **Background**: Recent studies show that community monitoring
improves public service outcomes in developing economies. Some of the papers
find that bottom-up accountability has a significant impact on public service
delivery:

	- Bjorkman, Martina and Jakob Svensson. 2009. “Power to the People: Evidence
		from a Randomized Field Experiment of a Community-Based Monitoring Project
		in Uganda.” Quarterly Journal of Economics 124(2):735–769.
	- Bjorkman, Martina and Jakob Svensson. 2010. “When Is Community-Based
		Monitoring Effective? Evidence from a Randomized Experiment in Primary
		Health in Uganda.” Journal of the European Economic Association
		8(2-3):571–581. Bjorkman Nyqvist, Martina, Damien de Walque and Jakob
		Svensson. 2017. “Experimental Evidence on the Long-Run Impact of
		Community-Based Monitoring.” American Economic Journal: Applied Economics
		9(1):33–69.

However, other authors fail to find positive resuls for the same type of
intervention, e.g.:

* Banerjee, A. V., Banerji, R., Duflo, E., Glennerster, R., and Khemani, S.
	2010. Pitfalls of Participatory Programs: Evidence from a Randomized
		 Evaluation in Education in India. American Economic Journal: Economic
		 Policy, 2(1):1–30.
* Lieberman, E. S., Posner, D. N., and Tsai, L. L. 2014. Does Information Lead
	to More Active Citizenship? Evidence from an Education Intervention in Rural
	Kenya. World Development, 60:69–83. Raffler, Pia, Dan Posner, and Doug
	Parkerson. 2019. The Weakness of Bottom-up Accountability: Experimental
	evidence from the Ugandan health sector. URL:
	[`https://bit.ly/2L36vzI`](http://egap.org/sites/default/files/Posner%2C%20Dan_Paper.pdf)

This project re-evaluates this theoretical controversy and proposes a new field
experiment to test whether bottom-up accountability is effective to promote
better policy outcomes. The results are substantively important as well. First,
we conduct this experiment in a middle-income country, Brazil, but which has
municipalities whose social indicators match both those of developed and
developing nations. Brazil is notably unequal in several characteristics, thus
allowing us to infer results for different subgroups if required. Second, we
are evaluating school constructions, and the presence of schools have essential
consequences in terms of long-run economic development, the household income
composition, and woman empowerment.

## Subjects and Context

**Eligibility and Exclusion Criteria**: We selected all school projects that
received Federal fundings from the Brazilian Ministry of Education. By an
agreement with the Brazilian Ministry of Education, they allowed us to have the
data of all school constructions receiving funds from them.

All schools constructions funded by the Brazilian Ministry of Education
participated in this study. We had no schools excluded from the program.

Interventions’ dates: We had three interventions here:

* App impact evaluation 1: From August 2017 to February 2018. We call this
	Intervention 1 throughout the text.
* App impact evaluation 2: From August 2018 to February 2019. We call this
	Intervention 2 throughout the text.

## Allocation Methods

**Assignment procedure**: The treatment here was assigned by the municipality
in the first intervention, and at the school level in the second and third
interventions.

For intervention one, we selected 150 municipalities that were placed in the
control group. The randomization code was:

```{r rand-int1, eval=FALSE}
vectreat <- c(rep(0,150), rep(1, nrow(dat)-150))

AIC_info = numeric()
for (i in 1:5000) {
  vectreat <- sample(vectreat, length(vectreat))
  mod <- glm(formula(paste('vectreat', vars, sep = '~')),
             family = binomial, data = dat)
  if(sum(summary(mod)$coefficients[-1,4]<.2)==0) {
    dat = data.frame(dat, vtreat = vectreat)
    AIC_info = c(AIC_info, AIC(mod))
  }
}
```

And then we checked whether the randomisation had one of the covariates
selected here significantly predicting the random assignment. We presented a
vector of acceptable random assignment in these grounds, and the Brazilian
Transparency selected the one we are using in the first intervention.

For intervention 2, we selected 15% of the schools in the dataset that could
either be placed in the treatment or in control in each of the interventions.
The `randomizr` code for intervention 2 follows below. `OnApp` means treatment
and `OffApp` means the control for this intervention:

```{r rand-int2, eval=FALSE}
# Install and load required package
install.packages("randomizr")
library(randomizr)
decl_intheapp <- declare_ra(blocks = dat_intheapp$var_blocking,
                            prob_each = c(.85,.15),
                            conditions = c('OnApp', 'OffApp'))

dat_intheapp$Z_intheapp <- block_ra(blocks = dat_intheapp$var_blocking,
                       conditions = c('OnApp', 'OffApp'),
                       prob_each = c(.85,.15))

dat_intheapp$IPW_intheapp <- 1/obtain_condition_probabilities(decl_intheapp,
                                                              dat_intheapp$Z_intheapp)
```

\noindent
**Block randomization**: We used simple randomisation in the first assignment.
In the second intervention, we blocked by the Brazilian State, the construction
status, and above the median execution. The summary of the blocks, with the
IPWs assigned follow below.

```{r block-int2, echo=TRUE, results='asis', cache=TRUE}
aux <- rand2018 %>%
  select(state, status, above_median_executed,
         var_blocking, Z_campaign, IPW_campaign) %>%
  unique() %>%
  na.omit()
kable(aux, "latex", caption = "Block randomization -- Intervention 2",
      booktabs = TRUE, longtable = TRUE) %>%
  kable_styling(font_size = 8) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"))
```

As pre-treatment variables, we used the following municipality-level
characteristics:

1. Log of Municipal Population in 2015
2. Log of Number of Poor Families (2010 IBGE Census)
3. Log of Total Federal Transfers to the Municipality in 2016
4. IDEB Indicator for primary school quality (2015 Ministry of Education)
5. IDEB Indicator for secondary school quality (2015 Ministry of Education)

All the pre-treatment variables were not significant, as displayed in Section
\ref{sec:bal-tests}.

## Treatments

* **Treatment groups**:
	- *Intervention 1*: Municipality with all schools funded by the Ministry of
		Education showing up in the app.
	- *Intervention 2*: School construction showing up in the app.

* **Control groups**:
	- *Intervention 1*: Municipality with all schools funded by the Ministry of
		Education **not** showing up in the app. 
	- *Intervention 2*: Selected school constructions **not** showing up in the
		app.

* **Method of manipulation delivery**:
	- *Intervention 1*: The municipalities in the treatment had all their school
		constructions showing up in the app. 
	- *Intervention 2*: The schools in the treatment group were showing up in the
		app.

* **Software**: The TDP app is an Android and iOS application developed to
	facilitate bottom-up pressure on school constructions in Brazil.

## Results

* **Outcome measures**: We use six outcome measures, all taken from the
	Ministry of Education biannual report:

1. Percentage of the investment executed before the impact evaluation started
	 (placebo)
2. Percentage of the investment executed by the end of the impact evaluation
	 period
3. The difference of the percentage invested in the end and in the beginning of
	 the impact evaluation period
4. Indicator for a construction finished during the impact evaluation period
5. Indicator for a construction canceled during the impact evaluation period
6. Indicator for an updated conclusion date for the construction during the
	 impact evaluation period

* **CONSORT**

Below follows the CONSORT chartflow:

  + *Intervention 1*:

```{r, echo=FALSE}
# Values ------------------------------------------------------------------
# Total dataset
ntotini <- dim(impact_evaluation_phase1)[1]

# Excluded
excld <- sum(is.na(impact_evaluation_phase1$treat))

# Total used randomization
diffrand <- ntotini-excld

# On treatment
tralloc <- sum(impact_evaluation_phase1$treat=='Treatment', na.rm = T)

# Control
ctrlalloc <- sum(impact_evaluation_phase1$treat=='Control', na.rm = T)

# Attrition treatment
lostfollowuptreat <- 0

# Attrition control
lostfollowupctrl <- 0

# Analyzed treatment
analtreat <- tralloc-lostfollowuptreat

# Analyzed control
analctrl <- ctrlalloc-lostfollowupctrl

values <- c(ntotini, excld, diffrand, tralloc, ctrlalloc,
            lostfollowuptreat, lostfollowupctrl, analtreat, analctrl)

# Defining Text Labels ----------------------------------------------------
text <- c('Assessment for\neligibility',
          'Excluded',
          'Randomized',
          'Allocated to\nTreatment',
          'Allocated to\nControl',
          'Lost to follow-up',
          'Lost to follow-up',
          'Analysed',
          'Analysed')

# Defining Function -------------------------------------------------------
paste1 <- function(x, y){
  paste0(x, ' (n=', y, ')')
}

# Concatenating Values and Text Labels ------------------------------------
LABS <- paste1(text, values)

ndf <-
  create_node_df(
    n = 21,
    label = c('Enrollment', 'Allocation', 'Follow-Up', 'Analysis',
              LABS, rep("", 8)),
    style = c(rep("solid", 13), rep('invis', 8)),
    shape = c(rep("plaintext", 4),
              rep("box", 9),
              rep("point", 8)),
    width = c(rep(2, 4), rep(2.5, 9), rep(0.001, 8)),
    hight = c(rep(0.5, 13), rep(0.001, 8)),
    fontsize = c(rep(14, 4), rep(10, 17)),
    fontname = c(rep('Arial Rounded MT Bold', 4), rep('Courier New', 17)),
    penwidth = 2.0,
    fixedsize = "true")

edf <-
  create_edge_df(
    arrowhead = c(rep('none', 3), rep("vee", 3), rep('none', 2), "vee", rep('none', 6),
                  rep("vee", 3), rep("none", 3), "vee", rep("none", 10)),
    color = c(rep('#00000000', 3), rep('black', 6), rep('#00000000', 6),
              rep('black', 3), rep('#00000000', 3), rep('black', 1),
              rep('#00000000', 2), rep('black', 2),
              rep('#00000000', 6)),
    constraint = c(rep("true", 18), rep('false', 14)),
    from = c(1, 19, 20, 16, 8, 10, # column 1
             5, 14, 7, 15, 2, 3, # column 2
             18, 6, 21, 17, 9, 11, # column 3
             1, 5, # row 1
             19, 14, # row 2
             20, 7, # row 3
             16, 15, # row 4
             8, 2, # row 5
             10, 3, # row 6
             12, 4), # row 7
    to = c(19, 20, 16, 8, 10, 12, # column 1
           14, 7, 15, 2, 3, 4, # column 2
           6, 21, 17, 9, 11, 13, # column 3
           5, 18, # row 1
           14, 6, # row 2
           7, 21, # row 3
           15, 17, # row 4
           2, 9, # row 5
           3, 11, # row 6
           4, 13)) # row 7

# Create Graph ------------------------------------------------------------
g <- create_graph(ndf,
                  edf,
                  attr_theme = NULL)

# Plotting ----------------------------------------------------------------
render_graph(g)
```

  + *Intervention 2*:

```{r, echo=FALSE}
# Values ------------------------------------------------------------------
# Total dataset
ntotini <- dim(impact_evaluation_phase2)[1]

# Excluded
excld <- sum(is.na(impact_evaluation_phase2$treat))

# Total used randomization
diffrand <- ntotini-excld

# On treatment
tralloc <- sum(impact_evaluation_phase2$treat=='Treatment', na.rm = T)

# Control
ctrlalloc <- sum(impact_evaluation_phase2$treat=='Control', na.rm = T)

# Attrition treatment
lostfollowuptreat <- 0

# Attrition control
lostfollowupctrl <- 0

# Analyzed treatment
analtreat <- tralloc-lostfollowuptreat

# Analyzed control
analctrl <- ctrlalloc-lostfollowupctrl

values <- c(ntotini, excld, diffrand, tralloc, ctrlalloc,
            lostfollowuptreat, lostfollowupctrl, analtreat, analctrl)

# Defining Text Labels ----------------------------------------------------
text <- c("Assessment for\neligibility",
          'Excluded',
          "Randomized",
          "Allocated to\nTreatment",
          "Allocated to\nControl",
          "Lost to follow-up",
          "Lost to follow-up",
          "Analysed",
          "Analysed")

# Defining Function -------------------------------------------------------
paste1 <- function(x, y){
  paste0(x, ' (n=', y, ')')
}

# Concatenating Values and Text Labels ------------------------------------
LABS <- paste1(text, values)

ndf <-
  create_node_df(
    n = 21,
    label = c('Enrollment', 'Allocation', 'Follow-Up', 'Analysis',
              LABS, rep("", 8)),
    style = c(rep("solid", 13), rep('invis', 8)),
    shape = c(rep("plaintext", 4),
              rep("box", 9),
              rep("point", 8)),
    width = c(rep(2, 4), rep(2.5, 9), rep(0.001, 8)),
    hight = c(rep(0.5, 13), rep(0.001, 8)),
    fontsize = c(rep(14, 4), rep(10, 17)),
    fontname = c(rep('Arial Rounded MT Bold', 4), rep('Courier New', 17)),
    penwidth = 2.0,
    fixedsize = "true")

edf <-
  create_edge_df(
    arrowhead = c(rep('none', 3), rep("vee", 3), rep('none', 2), "vee", rep('none', 6),
                  rep("vee", 3), rep("none", 3), "vee", rep("none", 10)),
    color = c(rep('#00000000', 3), rep('black', 6), rep('#00000000', 6),
              rep('black', 3), rep('#00000000', 3), rep('black', 1),
              rep('#00000000', 2), rep('black', 2),
              rep('#00000000', 6)),
    constraint = c(rep("true", 18), rep('false', 14)),
    from = c(1, 19, 20, 16, 8, 10, # column 1
             5, 14, 7, 15, 2, 3, # column 2
             18, 6, 21, 17, 9, 11, # column 3
             1, 5, # row 1
             19, 14, # row 2
             20, 7, # row 3
             16, 15, # row 4
             8, 2, # row 5
             10, 3, # row 6
             12, 4), # row 7
    to = c(19, 20, 16, 8, 10, 12, # column 1
           14, 7, 15, 2, 3, 4, # column 2
           6, 21, 17, 9, 11, 13, # column 3
           5, 18, # row 1
           14, 6, # row 2
           7, 21, # row 3
           15, 17, # row 4
           2, 9, # row 5
           3, 11, # row 6
           4, 13)) # row 7

# Create Graph ------------------------------------------------------------
g <- create_graph(ndf,
                  edf,
                  attr_theme = NULL)

# Plotting ----------------------------------------------------------------
render_graph(g)
```

* **Reasons for exclusion in the CONSORT:**

	+ *Intervention 1*: Constructions that were not in the app when the
		intervention started.

	+ *Intervention 2*: Constructions that were not in the app when the
		intervention started.

* **Statistical analysis**:

For all interventions, we run the following regression model:

$$ Y_i = \alpha + \beta T_i + \gamma X_i + \theta F_i + \varepsilon_i $$

Where $i$ indexes a given school observed in the intervention. $Y_i$ represents
an outcome variable. $T_i$ represents a treatment indicator. $\beta$ represents
the estimated Average Treatment Effect. $\gamma$ is a vector of pre-treatment
coefficient effects and $X_i$ a vector of pre-treatment covariates. $\theta$
represents a vector of fixed effects estimands and $F_i$ the Brazilian state
level fixed effects indicator vector. $\varepsilon_i$ is the common error term.

We run three types of analysis:

1. Full regressions with all data available
2. Full regressions using inverse probability weights
3. Regression where we match the control group with a same-size treatment group on the proximity of covariates using genetic matching.

All models use municipal-level cluster robust standard errors and state-level fixed effects. For all models, we also run regressions without clustering and without fixed effects.

# Other information

* **IRB**: FGV exempted the investigators from getting IRB approval because
	this research used online data and was approved by the Brazilian Ministry of
	Education.

* **Pre-registration**: this research has been pre-registered on the EGAP
	pre-registry tool: <https://egap.org/registration/4505>.

* **Replication materials**: The replication materials for this project are
	available here: <https://github.com/umbertomig/tdp-accountability>.

* **Funding**: This research received funding from the 2016 Google Social
	Impact Challenge. Google did not interfere in any aspect of the research
	design and analysis.

* **Conflict of interests**: The authors of this analysis declare that there
	are no conflict of interest.

# Session Information

We used `r R.version.string` to write the manuscript and this appendix. For the
regression models estimation, we used the `lfe` package. We employed the
`MatchIt` package to improve covariate matching. Everything in this report is
fully automated and can be reproduced using R Markdown.

```{r}
sessionInfo()
```

\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

# Bibliography